The Courtroom Challenge—A Conjecture on the Unavoidable Cost of LLM Verification

### 2025-08-03 paper/main.tex
Add files via upload
Tweaks to sketch proofs of lemmas.

\paragraph{Per-round notation.}
Fix a margin parameter $\delta\in(0,1]$.
For every run $\omega$ and round $t=1,\dots,T$ set
\begin{align*}
  y_t      &:=y_t(\theta)          &&\text{(ground-truth label)},\\
  h_t      &:=h_t(\omega)          &&\text{(validator's automatic prediction)},\\
  \mathrm{Esc}_t&:=\mathbf 1_{\{\text{round $t$ is escalated}\}.}
\end{align*}
Define the \emph{leverage rounds}
\begin{equation}\label{eq:kappaDef}
  \mathcal L(\omega):=\Bigl\{\,t:\,
     \mathrm{Esc}_t=0\;\text{and}\;|y_t-h_t|\ge\delta\Bigr\},
  \qquad
  \kappa(\omega):=\frac{|\mathcal L(\omega)|}{T}.
\end{equation}
The cumulative label error on those rounds is
\begin{equation}\label{eq:ErrDef}
  \Err(\omega):=\sum_{t\in\mathcal L(\omega)}
                \mathbf 1_{\!\bigl[h_t\neq y_t\bigr]}.
\end{equation}
Define the global means:
\[
  \bar\kappa:=\E[\kappa],
  \qquad
  \Err_{\mathrm{bar}}:=\E[\Err].
\]

@@

\begin{proof}[Proof sketch]
Step 1 (separation).
For any $m\neq m'$ and transcript $x$ at most one of
$y_t(m),y_t(m')$ can equal $h_t$ on a leverage round, hence
\[
  \Err_m(x)+\Err_{m'}(x)\;\ge\;\delta\,\kappa(\omega)\,T.
  \tag{$\ast$}
\]

@@

\begin{proof}[Proof sketch]
The rule ensures $M_T$ is a $\tau$-packing of $\mathcal Q$.
By a standard $\varepsilon$-net/volume argument, any $\tau$-packing inside a
$D$-diameter set has size at most $(D/\tau)^{d}$.
\end{proof}

@@

\begin{proof}[Sketch]
1.~Maximal $\tau$-packing.
Let $\{c_i\}_{i=1}^{m}\subset\mathcal Q$ be maximal with pairwise distances
$>\tau$.  By inscribing disjoint axis-aligned cubes of side $\tau/\sqrt d$ we
obtain $m\ge\alpha_d(D/\tau)^d$ with $\alpha_d=2^{-d}$.

\noindent
2.~Disjoint balls.
Set $B_i=B(c_i,\tfrac{\tau}{2})$.  The balls are disjoint and lie inside
$\mathcal Q$.

@@

\noindent
5.~Union bound and insertion rule.
With probability $\ge1-\delta$ every $B_i$ is hit at least once.
The first hit in each ball is $>\tau$ from all stored points,
hence inserted; so $|M_T|\ge m\ge\alpha_d(D/\tau)^d$.
Taking $c_d=\tfrac12\nu_d^{-1}D^{-d}\le\alpha_d$ matches the lemma's constant.
\end{proof}

>>>>

\paragraph{Per-round notation.}
Fix a margin parameter $\delta\in(0,1]$.
For every run $\omega$ and round $t=1,\dots,T$ set
\begin{align*}
  y_t      &:=y_t(\theta)          &&\text{(ground-truth label)},\\
  h_t      &:=h_t(\omega)          &&\text{(validator's automatic prediction)},\\
  \mathrm{Esc}_t&:=\mathbf 1_{\{\text{round $t$ is escalated}\}}.
\end{align*}
Define the \emph{leverage rounds}
\begin{equation}\label{eq:kappaDef}
  \mathcal L(\omega):=\Bigl\{\,t:\,
     \mathrm{Esc}_t=0\;\text{and}\;|y_t-h_t|\ge\delta\Bigr\},
  \qquad
  \kappa(\omega):=\frac{|\mathcal L(\omega)|}{T}.
\end{equation}
The cumulative label error on those rounds is
\begin{equation}\label{eq:ErrDef}
  \Err(\omega):=\sum_{t\in\mathcal L(\omega)}
                \mathbf 1_{\!\bigl[h_t\neq y_t\bigr]}.
\end{equation}
For any hypothetical label index \(m\) and transcript \(x\) we write
\[
  \Err_m(x):=\sum_{t\in\mathcal L(\omega)}\mathbf 1_{\!\bigl[h_t(x)\neq y_t(m)\bigr]},
\]
i.e.\ the cumulative error that would occur on the same leverage rounds if the
hidden parameter were \(m\).
Define the global means:
\[
  \bar\kappa:=\E[\kappa],
  \qquad
  \Err_{\mathrm{bar}}:=\E[\Err].
\]

@@

\begin{proof}[Proof sketch]
Step 1 (separation).
Fix \(m\neq m'\) and transcript \(x\).  On each leverage round
\(t\in\mathcal L(\omega)\) we have \(|y_t(\theta)-h_t|\ge\delta\).
Hence \(h_t\) can coincide with at most one of
\(y_t(m),y_t(m')\), so
\[
  \Err_m(x)+\Err_{m'}(x)\;\ge\;\delta\,\kappa(\omega)\,T\;.
\tag{$\ast$}
\]

@@

\begin{proof}[Proof sketch]
The rule ensures $M_T$ is a $\tau$-packing of $\mathcal Q$.
By a standard $\varepsilon$-net/volume argument, any $\tau$-packing inside a
$D$-diameter set has size at most $(D/\tau)^{d}$.
(In the worst‑case geometry one needs an extra factor \(2^{d}\); we omit it
because all subsequent results depend only on the \(\tau^{-d}\) scaling.)
\end{proof}

@@

\begin{proof}[Sketch]
1.~Maximal \(3\tau\)-packing.\footnote{Using \(3\tau\) rather than \(\tau\) only
shrinks the packing constant by \(3^{-d}\); the \(\Theta(\tau^{-d})\) scaling
is unaffected.}
Let \(\{c_i\}_{i=1}^{m}\subset\mathcal Q\) be maximal with pairwise
distances \(>3\tau\).  Axis‑aligned cubes of side \(3\tau/\sqrt d\) give
\(m\ge\alpha_d(D/\tau)^d\) with \(\alpha_d=3^{-d}2^{-d}\).

\noindent
2.~Disjoint balls and guaranteed insertion.  
Set \(B_i=B(c_i,\tfrac{\tau}{2})\).  
For any query \(q_t\in B_i\) and any prior memory point \(m\in B_j\;(j\neq i)\),
\[
\lVert q_t-m\rVert
\;>\; 3\tau-\tfrac{\tau}{2}-\tfrac{\tau}{2}
\;=\;2\tau>\tau,
\]
so the insert-iff-distance \(>\tau\) rule is automatically satisfied.
The
balls are disjoint and lie inside \(\mathcal Q\).

@@

\noindent
5.~Union bound and memory lower bound.  
With probability \(\ge 1-\delta\) every \(B_i\) is hit at least once.
If the first hit is accepted, it is written to memory; if it is rejected, the
blocking memory point \(m\) satisfies
\(\lVert m-c_i\rVert<1.5\,\tau\) and therefore lies in the expanded ball
\(\tilde{B}_i:=B(c_i,1.5\,\tau)\).
Because the \(\tilde{B}_i\) are disjoint (centre separation \(>3\tau\)),
each centre still accounts for one distinct stored point, so
\(|M_T|\ge m\ge\alpha_d(D/\tau)^d\).
\end{proof}
